### 2017-02-20 case.tex

\section{Case Studies}
\label{sec:case}

We present case studies on applying our validation framework on two protocols
implemented in the \ns{} network simulator: 802.11 data transmission and ARF
rate control algorithm. The goal is to demonstrate how our framework can avoid
false alarms and report true violations on incomplete sniffer traces and report
true violations. We also report the application of our framework to a commercial
product under development.

\input{./case/tx.tex}
\input{./case/arf.tex}
\input{./case/xbox.tex}