### 2017-02-20 conclusion.tex
conclusion

\section{Conclusions}
\label{sec:conclusion}

We formally define the uncertainty problem in validating wireless protocol
implementations using sniffers. We describe a systematic augmentation of the
protocol state machine to explicitly encode the uncertainty of sniffer traces.
We characterize the NP-completeness of the problem and propose both an
exhaustive search algorithm and heuristics to restrict the search to more likely
traces. We present two case studies using \ns{} network simulator to demonstrate
how our framework can improve validation precision and detect real bugs. We also
report the application of our framework on a commercial product under
development.

Finally, we discuss a few challenges and future
directions.

\textbf{Verification Coverage.} Given a single sniffer trace, it is possible
that not all the states in the state machine are visited during the verification
process. For instance, a rate control state machine based on certain consecutive
packet losses patterns can not be verified if no such consecutive losses appear
in the sniffer trace. In general, given a protocol state machine, it is
challenging to extract the packet patterns for each state to be reached and to
alter the testing such that such patterns can be observed.

\textbf{State Machine Generation.} We manually translated the protocols studied
in this paper into monitor state machines based on the source code, comments and
documentation. The process is time-consuming and error-prone. A more scalable
approach would be taking the protocol specification written in certain formal
language, and automatically translate such specification into state machines
that can be used for verification process.