### 2017-07-09 case/tx.tex
comments

\subsection{802.11 Data Transmission}
\label{subsec:tx}

In this section, we first show that our framework can improve validation
precision by inferring the missing or extra packets using the augmented
transition framework. We then demonstrate the ability of our framework to detect
true violations by manually introducing bugs in the \ns{} implementation and
show the precision and recall of validation results.

\subsubsection{Experimental Setup}

We set up two \wifi{} devices acting as the transmitter (DUT) and receiver
respectively. Another \wifi{} device is configured in monitor mode and acts as
the sniffer. During the experiments,we collect both the DUT packet trace
(the ground truth) and the sniffer trace.

The DUT and endpoint are configured to use the IEEE 802.11g standard with both
RTS/CTS and fragmentation disabled.  A Constant Bit Rate (CBR) UDP traffic
(54~Mbps) is generated from the DUT to the endpoint.  The UDP packet size is
1436 bytes, which results in a 1500 bytes \wifi{} packet.

In order to control the packet loss ratios between each pair of devices, we
developed a new propagation loss model for \ns{} called
\texttt{MatrixRandomPropagationLossModel}.
\sloppy{%
  Instead of a constant propagation
  loss as in existing \texttt{MatrixPropagationLossModel}, the signal
  propagation loss between a pair of nodes is determined by a binary random
  variable of two  values: 0~dB (no loss) and 1000~dB (complete loss).
  }
Therefore, any packet loss
probability can be achieved by adjusting the random variable distribution.
The
model supports both symmetric and asymmetric propagation losses.
We use
symmetric propagation loss in all our experiments.
Finally, pcap capture is enabled in both the DUT and the sniffer devices.

\subsubsection{Verifying Unmodified Implementation}

In the original monitor state machine shown in Fig.~\ref{fig:dot11_tx_ta}, we
set acknowledgment timeout $T_o=334\mu$s, maximum retransmission delay
$T_m=15$ms according to the protocol. We also adapt the state machine to include
multiple retransmissions\footnote{The exact number of retransmissions is not
part of the protocol, and NS-3 implementation set this to be 7.} instead of one.

Let $Pr_{ds}$, $Pr_{es}$ and $Pr_{ed}$ be the packet loss probability between
the DUT and sniffer, endpoint and sniffer, endpoint and DUT respectively.
$Pr_{ed}$ represents the characteristics of the system being tested, while
$Pr_{ds}$ and $Pr_{es}$ represent the sniffer's quality in capturing packets.

We vary each of the three probabilities, $Pr_{ds}$, $Pr_{es}$ and $Pr_{ed}$, from
0 to 0.5 (both inclusive) with 0.05 step.  For each loss ratio combination, we
ran the experiment 5 times, and each run lasted
30 seconds. In total, 6655 ($11^3\times 5$) pairs of DUT and sniffer packet
traces were collected.

To establish the ground truth of violations, we first verify the DUT packet
traces using the \textit{original} state machine $S$.  This can be achieved by
disabling augmented transitions in our framework.  As expected, no violation is
detected in any DUT packet traces.

We then verify the sniffer traces using the augmented state machine $S^+$.  For
the $\mathit{GoBack}(k)$ heuristic, we set $k=7$, which is the maximum number of
transmissions of a single packet. For the $\mathit{NumMissing}(d, l, k)$ heuristic, we
set the sliding window size $l=100$, and $k=80$ such that no violation is
reported. The relationship of $k$ and validation precision is studied in next
section.

Next, we present detailed analysis of the augmented transitions on the sniffer
traces. The goal is to study for a given system packet loss probability
$Pr_{ed}$, how the sniffer packet loss properties ($Pr_{ds}$ and $Pr_{es}$)
affect the difference between the DUT trace and the mutation trace, which represents
a guess of the DUT trace by the augmented state machine based on the sniffer
trace.

For all following analysis, we divide the traces into three groups according to
$Pr_{ed}$: low ($0 \le Pr_{ed} \le 0.15$), medium ($0.20 \le Pr_{ed} \le 0.35$)
and high ($0.40 \le Pr_{ed} \le 0.50$).

The difference between two packet traces can be quantified by the Jaccard distance
metric.

\begin{align}
  Jaccard(Tr_1, Tr_2) = \frac{\left\vert Tr_1 \ominus Tr_2\right\vert}{\left\vert
  Tr_1 \cup Tr_2\right\vert}
\end{align}%
where $\ominus$ is the symmetric difference operator.  The distance is 0 if the
two traces are identical, and is 1 when the two traces are completely different.
The smaller the distance is, the more similar the two traces are.

A na\"ive way to calculate the Jaccard distance of two traces is to use the hash
of the $(time, packet)$ pair for set intersection and union operation.  However,
it does not work for mutation trace which contains fabricated packets with no
actual payload.  Therefore, we use a protocol specific canonical representation
of packets when calculating the distance.  In particular, the string
\textit{r\_DATA\_i\_t} represents the $t^{th}$ transmission of a data packet
with sequence number $i$, and $r$ represents the round of sequence numbers as it
wraps after 4096.  And similarly \textit{r\_ACK\_i\_t} is the corresponding
acknowledgment packet.

\begin{figure*}[t!]
  \centering
  \begin{subfigure}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{MutationDUTJaccard3DFigure_0_10.pdf}
    % \caption{$Pr_{ed} \in [0.05, 0.1, 0.15]$}
    \caption{$0.05 \le Pr_{ed} \le 0.15$}
  \end{subfigure}\hspace*{0.01\textwidth}
  \begin{subfigure}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{MutationDUTJaccard3DFigure_0_30.pdf}
    \caption{$0.2 \le Pr_{ed} \le 0.35$}
  \end{subfigure}\hspace*{0.01\textwidth}
  \begin{subfigure}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{MutationDUTJaccard3DFigure_0_50.pdf}
    \caption{$0.4 \le Pr_{ed} \le 0.5$}
  \end{subfigure}\hspace*{0.01\textwidth}
  \caption{\textbf{Jaccard Distance Between Mutation and DUT Traces.} For each
  data point, the mean of the 5 runs is used.}
  \label{fig:mutation_dut}
\end{figure*}

Fig.~\ref{fig:mutation_dut} shows the Jaccard Distance between mutation and its
corresponding DUT trace. We make the following observations. First, for a given
system loss probability $Pr_{ed}$ (each sub-figure), the lower the sniffer
packet loss probability ($Pr_{ds}$ and $Pr_{es}$), the smaller Jaccard distance
between the DUT and mutation trace. Intuitively, this means a sniffer that
misses less packets can enable our framework to better reconstruct the DUT
trace.

Second, we observe a \textit{protocol-specific} trend that $Pr_{ds}$ is more
dominant than $Pr_{es}$.  This is because retransmission packets of the same
sequence number are identical, hence when the sniffer misses multiple
retransmission packets, our framework only needs to infer one retransmission
packet to continue state machine execution.

Finally, as the system loss probability $Pr_{ed}$ increases, the Jaccard
distance increases more rapidly as $Pr_{ds}$ increases.  This is because the
ratio of retransmission packet increases along with $Pr_{ed}$.

We then evaluate the cost of resolving uncertainty. In particular, we use
the Average Search Steps Per Packet (ASSPP) as a metric to quantify the search
cost.  It is calculated by dividing the total number of search steps by the
number of packets in the packet trace. For DUT traces, ASSPP is always 1 since
there is no uncertainty. For sniffer traces, however, multiple search steps must
be conducted to resolve the potential uncertainty of each packet in the
sniffer trace.

\begin{figure*}[t!]
  \centering
  \begin{subfigure}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{StepCount3DFigure_0_10.pdf}
    \caption{$Pr_{ed} \in [0.05, 0.10, 0.15]$}
  \end{subfigure}\hspace*{0.01\textwidth}
  \begin{subfigure}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{StepCount3DFigure_0_30.pdf}
    \caption{$Pr_{ed} \in [0.2, 0.25, 0.3, 0.35]$}
  \end{subfigure}\hspace*{0.01\textwidth}
  \begin{subfigure}{0.33\textwidth}
    \centering
    \includegraphics[width=\textwidth]{StepCount3DFigure_0_50.pdf}
    \caption{$Pr_{ed} \in [0.40, 0.45, 0.50]$}
  \end{subfigure}
  \caption{\textbf{Average Searching Step Per Packet.} For each data point, the mean of 5 runs
  is used.}
  \label{fig:cost}
\end{figure*}

Fig.~\ref{fig:cost} shows the distribution of ASSPP at different $Pr_{ed}$.
Similar to the case in Fig.~\ref{fig:mutation_dut}, $Pr_{ds}$ plays a dominant
role in determine the searching cost. One interesting observation for this
particular protocol is that the
search cost peaks when $Pr_{ds}$ is high while $Pr_{es}$ is low. In such loss
probability combinations, sniffer misses many data packets from the DUT but
picks up lots of \textit{dangling} $Ack$ packets from the DUT. Because the $Ack$
packet has neither sequence numbers nor retry flag, the searching algorithm had
a hard time resolving such uncertainty.

\subsubsection{Introducing Bugs}

We have demonstrated that our framework can tolerate sniffer imperfections and
avoid raising false alarms.  The next question is, can it detect true
violations?  To answer this question, we manually introduce several bugs in
\ns{} implementation that concerns various aspects of 802.11 data transmission
protocol.  More specifically, the bugs are:

\begin{itemize}
  \item \textbf{Sequence Number}: the DUT does not assign sequence number
    correctly. For example, it may increase sequence by 2 instead of 1, or it
    does not increase sequence number after certain packet, etc. We choose one
    type of such bugs in each run.
  \item \textbf{Semantic}: the DUT may retransmit even
    after receiving $Ack$, or does not retransmit when not receiving $Ack$.
\end{itemize}

We instrument the \ns{} implementation to embed instances of bugs in each
category. At each experiment run, we randomly decide whether and which bug to
introduce for each category. We fix $Pr_{ds}=Pr_{es}=0.1$ and vary $Pr_{ed}$
from 0.0 to 0.5 with 0.01 step. For each $Pr_{ed}$ value, we ran the experiment
100 times, of which roughly 75 experiments contained bugs. In total, 5100 pairs of
DUT and sniffer traces were collected.

We use the DUT packet traces as ground truth of whether or not each experiment
run contains bugs.
For each $Pr_{ed}$ value, we calculate the precision and recall of violation
detection using the sniffer traces.%
\begin{align}
  \text{Precision} = \frac{\left\vert \{\text{Reported Bugs}\} \cap \{\text{True Bugs}\}\right\vert}{\left\vert
  \{\text{Reported Bugs}\}\right\vert}\\
  \text{Recall} = \frac{\left\vert \{\text{Reported Bugs}\} \cap \{\text{True Bugs}\}\right\vert}{\left\vert
  \{\text{True Bugs}\}\right\vert}
\end{align}%
The precision metric quantifies how \textit{useful} the validation results are ,
while the recall metric measures how \textit{complete} the validation results
are.

\begin{figure}[t!]
  \centering
  \begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\textwidth]{PrecisionFigure.pdf}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
    \includegraphics[width=\textwidth]{RecallFigure.pdf}
  \end{subfigure}
  \caption{\textbf{Precision and Recall of Validation Results.}}
  \label{fig:precision}
\end{figure}

Fig.~\ref{fig:precision} shows the CDF of precision and recall of the 51
experiments for various $k$ values. For precision, as expected, the more
tolerant the search to sniffer losses (larger $k$), the more tolerant the
framework is to sniffer losses, and the more precise the violation detection. In
particular, when $k=30$, the precisions are 100\% for all $Pr_{ed}$ values.
Second, the recall is less sensitive to the choice of $k$.  Except for the
extreme case when $k=30$, all other thresholds can report almost all the
violations.