### 2017-07-09 submitted/response.txt
add response

First, we would like to thank all reviewers for their insightful comments and
suggestions. Besides fixing typos and minor cosmetic issues, we also tried our
best to address the comments. Responses to each individual reviewers are listed
below.

To reviewer 1:

With respect to the edit distance, that's precisely the idea behind augmented
transitions. Type-1 transitions *inserts* packets to the sniffer trace while
Type-2 transitions *removes* packets from the sniffer trace. The number of
augmented transitions (Type-1 and Type-2) represents the "distance" between the
mutation trace and the original sniffer trace. The gist of the NumMissing
heuristic we proposed is exactly to limit the "distance" (hence edits) to the
sniffer trace.  In other words, the "distance" between the mutation trace and
sniffer trace represents the confidence of protocol correctness.

To reviewer 2:

With respect to the verification coverage future work, the current proposal
provides ways to analyze sniffer traces that are already being collected. While
this is useful in revisiting existing traces which are mostly likely only spot
checked, the verification is still limited to a static sets of sniffer trace
which may only cover the "normal" path of a protocol excitation. Consider the
case study of the game controller protocol, the rate control related violations
were only detected in a tiny fraction of the traces since the violation requires
non-trivial packet loss patterns to manifest itself. The idea behind our
verification coverage proposal is to proactively control the process of
experiment set up and sniffer trace collection to increase the chance of each
protocol state being visited, hence increasing the  probability of violation
detection.

To reviewer 3:

With respect to the fault model, in this paper we only consider full packet
loss, which reflects most practical sniffer's behavior. In most wireless
communications protocols, per-packet checksum is used and a packet that fails
the checksum will be dropped by the network driver.

With respect to the detailed NS-3 experiment setup, we added a paragraph in
Section 5.1.1 to explain the packet loss model we used in the experiments.
Basically, we added a new propagation loss component to NS-3 so that we can
control per-packet losses between each individual pair of devices.